Generation of TypeScript Declaration Files from JavaScript Code

### 2021-05-26 mplr-2021/programming21-reviews-answered.txt

Said that for Globtoregexp.I__opts already!

> Fig 3: Why not depict the example extraction step --- which sometimes
> includes manual addition of examples --- here? And consider
> representing Jalangi's role in this architecture.

Fernando?
F: Added new images:
  - mplr-2021/graphics/dts-generate-block-diagram-with-code-examples-block.pdf
  - mplr-2021/graphics/dts-generate-block-diagram-with-jalangi.pdf

> Sec 4.2/4.3: As mentioned, I think a lot more detail is warranted
> about the instrumentation and the process for computing types from
> traces.
> 
> "Execution of Code Examples": Interesting to see how many examples
> were perhaps out-of-date or were not proper examples. Were there any
> interesting patterns worth discussing?
> 
> p.17: It's encouraging to see how the proposed tool can fix some
> errors found in manual type declarations!
> 
> "Literal types vs string": If considering asking developers to mark
> strings, instead why not ask them to directly use, e.g. the "Enums" or
> "Discriminated Unions" (https://www.typescriptlang.org/docs/handbook)?

Misconception! They write JavaScript, not TypeScript! Added a
paragraph of clarification.

> "Any": Ditto above; why not ask developers to choose the any type?
> 
> "Further types": Inferring indexed types, dependent types, etc. from
> run-time traces would seem to require quite significant technical
> extensions to the approach described here.
> 
> TSInfer: Is it fundamental that TSInfer apparently generates
> signatures for helpers? Could there not be a simple syntactic analysis
> that identifies top-level definitions, to be include in a declaration
> file?
> 
> 
> *** Minor Comments and Typos ***
> 
> Fig 1: Maybe I'm missing it, but I don't see Footnote 6 anywhere.j

FIXED

> p.5: "globToRegexp" ==> "GlobToRegexp" twice

FIXED

> p.8: I'm not sure why "syntax highlighting is done correctly" is
> noteworthy here.

REMOVED

> Fig 4: Is this referenced in the text?

FIXED

> p.15: "loose" ==> "lose"
>

FIXED

@@

> 
> ## Major comments
> 
> p. 1: I have the feeling that the abstract was not prepared in
> accordance to the specifications of the journal including the
> keywords context, inquiry, approach, knowledge, grounding and importance.
> It's not that I do not get an impression of the work presented by
> the paper, but it just feels like an "ordinary abstract" and
> not one that explicitly tackles the key questions behing the keywords above.

@@

> 
> p. 4: "We examined all 6029 entries in the DefinitelyTyped
> repository and found 244 sufficiently well documented NPM
> packages," - it's a shame that the final portion of repositories that can be tested is less than 5% of the starting value.

@@

> 
> p. 15: "In this step, we loose another 50% of modules! This loss is
> mainly explained because some developers do not wrap
> their code in a block using the javascript or js tags. However, as
> we are still left with code examples for 2260 modules, we did
> not further look into code extraction as this number was considered
> sufficient for evaluating the generation of declaration
> files." - Given the low value for the final number of repositories,
> you should definitely reconsider adapting your extraction
> mechanism to be more robust!

@@

> 
> p. 17: "We were unable to generate a declaration file based on the
> module-class template. All modules corresponding to this template
> exhibited at least one of the unimplemented TypeScript features." -
> I was wondering if could make sense to adapt
> the tooling machinery to leave out only the definitions that consist
> of unimplemented features, instead of leaving out a
> whole module if one of its definition.

@@

> 
> p. 18: "The remaining 174 files contained at least one of the
> unimplemented TypeScript features so that they could not be
> considered." - This sentence sounds like you indeed generate as much
> as possible and filter out only layer --- how does this
> statement relate my previous concern?

@@

> 
> ## Minor comments
> 
> p. 5: "The choice of the template depends on the structure of the
> underlying JavaScript library:..." - here the following
> listing/itemize environment seems to be missing, at least, I think that a itemize would be the better layout to use here.

> p. 14: "dts-compare does not contemplate differences in function
> return types, because interface inference of function
> return types is not covered in the current version of dts-generate."
> - Why is it that the inference for return types is that
> different from parameters? You infer the parameters through example
> code, by running example code, you get concrete return
> values that you can analyse in the same fashion. Furthermore, Figure
> 4 (for example) shows a generated file with functions
> with return types. What am I missing here?

Some explanation provided.

@@

> 
> p. 1: "..., which is error prone and time consuming." - error-prone
> p. 2: "..., which is time consuming and error prone." - error-prone
FIXED^2
> p. 8: "A TypeScript Declaration File" - declaration file (why
> title-case?)
FIXED

@@

> p. 8: "... so that test cases were difficult to reap from the NPM packages:..." - read
NOT FIXED
> p. 9: "The Runtime Information block in Figure 3" - run-time information (why title-case?)
FIXED
> p. 9: "For example, Function f ... Property foo ... Parameter a
> ... - function ... property ... parameter (it seems weird to start
> with upper-case letters for the itemization)

@@

> p. 11: " and traverse the Abstract Syntax Tree of a TypeScript program" - abstract syntax tree
OK
> p. 12: "DefinitelyTyped Repository" - repository (why title-case?)
OK
> p. 12: "uses the TypeScript Compiler API" - TypeScript compiler API
> (why title-case?)
OK

@@

> p. 18: "Excerpt of JS implementation that throws an Error when in- put parameter is not a string" - error
OK
> p. 19: "Properties of interface Result were ignored for readability"
> - result
NOT FIXED

@@

> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> The paper presents a tool that automatically generates TypeScript
> definition files for JavaScript projects. The tool uses > dynamic
> analysis. It runs the code samples found in the README file in the
> project repository, records information about > method invocations
> and argument values and uses those to generate type definitions. The
> paper evaluates the tool by > comparing generated definitions with
> hand-written definitions from the DefinitelyTyped repository - this
> is done for libraries > which have code samples in README,
> i.e. about ~270 out of ~6000. 

@@

> 
> The paper presents a tool that is useful, although it has a fairly
> limited scope. The approach presented by the authors - to > run a
> code sample using a library and use runtime monitoring to deduce the
> structure of the library - is definitely a promising > one. The
> authors make a good first step towards building a practical
> tool. However, the current implementation of the tool is > quite
> basic in a number of ways. First, it relies on code samples in
> README, which is not exactly common (it will work only > for quite
> basic libraries). Second, it ignores many interesting types that a
> library could use - most prominently arrays, but > those are just
> one example. 

@@

> DETAILED COMMENTS
> 
> * In the introduction, you say that manually writing definitions is error-prone, but your paper then uses manually written > definitions from DefinitelyTyped as ground truth. I find this a bit paradoxical - maybe you could instead say that this is a lot > of work and there are many libraries that are not covered by DefinitelyTyped?

ADDED a SENTENCE

> 
> * I have to admit that I was a bit confused by the idea of "templates" that is used in the paper. Do the templates determine > just how you structure the definition, or are they somehow limiting what kind of constructs can be used in a given file? I think > this would benefit from some clarification in the paper.

EXPLAINED in section 3

@@

> * There is very little detail in the paper about how you infer types from the sample values that are observed during the > runtime execution. I guess the mechanism here is quite basic, but it is still worth discussing. (You might be interested in the > related paper [1] which does this for more complex types, but also covers things like records with optional fields.)

DONE in section 4

> * I think your evaluation is somewhat limited by the fact that the libraries that have a complete documentation in the > README are the ones with very simple API. It would be interesting to evaluate how your approach works on larger libraries > (even if that meant copying some examples into a readme manually).
> 
> * I think the phrasing that "more advanced types are overblown" is a bit unfortunate, especially if the tool does not support > some quite basic types like arrays.

TONED DOWN. We have arrays now.

> * The paper should give a much more extensive overview of related work - for example, it should discuss what has been > done using static analyses, whether similar tools exist for other languages and there might also be other interesting related > approaches (perhaps some based on machine learning methods). The references in the submission are fairly minimal.
> 
> [1] T. Petricek, G. Guera, D. Syme. Types from data: making structured data first-class citizens in F#, PLDI 2016

### 2021-05-26 mplr-2021/main.tex
tightening up

\input{macros}
\newcommand{\secref}[1]{Section~\ref{#1} - \nameref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\coderef}[1]{Listing~\ref{#1}}

@@

  \caption{Declaration file for \texttt{glob-to-regexp} 
    generated with \texttt{dts-generate}} 
  \label{fig:motivating-example-glob-to-regexp-vscode}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
  \caption[dts-generate - Architecture overview]{dts-generate - Architecture
      overview - The initial code base is retrieved from the \NPM{} package's
    repository.
    Run-time information is gathered driven by example code.
    A TypeScript declaration file is generated using run-time
    information.
    % A Symbolic Execution Engine creates test cases based on the generated
    % Declaration File and via a feedback loop enriches the code base until the stopping
    % criteria is reached.
    % The final TypeScript Declaration File gets returned.
    % Feedback
    % loop through the Symbolic Execution Engine was not implemented. It can be added in a
    % future to the existing architecture, without modifying the working blocks.
  } 
  \label{fig:tsd_generation_method_block_diagram}
\end{figure}

@@

\begin{figure}[tp]
  \begin{centering}
      {\includegraphics[width=1\linewidth]{evaluation-diagram.pdf}}
      \caption[Evaluation against DefinitelyTyped repository]{\textbf{Evaluation of
          generated declaration files against DefinitelyTyped repository} - The parser
        uses the TypeScript Compiler           API \cite{typescript-compiler-api} to
        transform declaration files into an abstract syntax which is serialized in
        JSON. The comparison is performed on the JSON format.} 
      \label{fig:evaluation-diagram}
  \end{centering}
\end{figure}

@@

\begin{lstlisting}[
  caption={Example for \texttt{dts-compare} executed on declaration files without differences},
  language=bash,
label=code:dts-compare-example,
  float=tp,captionpos=b
]
$ dts-compare -e expected.d.ts -a actual.d.ts --module-name "my-module"
{
    "module": "my-module",
    "template": "module",
    "differences": [],
    "tags": []
}
\end{lstlisting}
% $

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=1\linewidth]{funnel.png}
  \caption[Number of analyzed modules for each stage of the experiment]{
    \textbf{Number of analyzed modules for each stage of the experiment} - 
  A TypeScript declaration file was generated for only \CountModulesGeneratedDeclarationFile{} modules, out of \CountTotalModulesDefinitelyTyped{} modules in the DefinitelyTyped repository. 
  % It was only possible to gather valid run-time information for about 25\% of the modules for which a Code Example was extracted.
  } 
  \label{fig:experiments-overall-funnel}
\end{figure}

@@

The results obtained for each step are described in the
following sections.

@@

\paragraph*{Generated Declaration Files}
A declaration file was generated for \CountModulesGeneratedDeclarationFile{} out of \CountModulesRunTimeInfoExtracted{} modules. Despite the correct execution
of the instrumented code examples, \CountModulesDeclarationFileCouldNotBeGenerated{} modules did not yield suitable run-time information
for generating a declaration file. The extracted code examples for these modules did not
execute the library. Hence, the collected run-time information was not sufficient
for generating a declaration file.

@@

\paragraph*{Literal types vs string}
Many JavaScript libraries use literal strings to select
options and configurations. To be concrete, 1925 DT packages rely on
literal types (see Table~\ref{tab:dts-parse-stats}). For example, the bonjour
package\footnote{\url{https://www.npmjs.com/package/@types/bonjour}}
relies on strings to control events and indicate protocols in (different) interfaces:
\begin{lstlisting}[language=TypeScript]
    removeAllListeners(event?: 'up' | 'down'): this;
    protocol?: 'udp'|'tcp';
\end{lstlisting}
The carlo
package\footnote{\url{https://www.npmjs.com/package/@types/carlo}}
uses literals to distinguish different 
types of events.
\begin{lstlisting}[language=TypeScript]
export type AppEvent = 'exit' | 'window';
\end{lstlisting}
% coinbase ...

@@

\section{Related Work}
\label{sec:related-work}
\paragraph*{Dynamic analysis techniques} Trace Typing
\cite{DBLP:conf/ecoop/AndreasenGCSTS16} is a framework for evaluating
retrofitted type systems. It involves gathering traces of JavaScript program
executions, just like in our system. However, the goal of the type
systems that they study is quite different from ours. They are
interested in type information to improve run-time performance and
traditional compilation, giving information about object layout and
tag tests. Contrary to their approach, we do not just observe the
types of arguments and return values. In addition, we also trace
property accesses which translate to requirements on objects. Also, we
have no generalization step to infer polymorphic types.

Hummingbird \cite{DBLP:conf/pldi/RenF16} is a system for type checking
Ruby programs at run time in the presence of
metaprogramming. Metaprograms are extended to additionally generate
type annotations for the generated code. Once a generated method is
called, the method's body is statically checked against the actual
argument types as well as the generated annotations. This system does
not involve run-time tracing.

@@

JSTrace \cite{saftoiu21:_jstrac} also performs dynamic type inference
for JavaScript by gathering information at run time. It inspects the
run-time types of arguments and return values of suitably wrapped
functions. While the resulting information is treated similarly as in
our work, we also gather interactions that result from property
accesses and method calls and derive further typing constraints from
those. Moreover, JSTrace performs an (expensive) deep inspection of
the types of objects and arrays whereas our inspection is shallow and
thus more efficient.

@@

As TSInfer analyzes the objects and functions stored in the snapshot, it faces the problem
of including internal methods and private properties in the declaration file. Run-time
information would have informed that the developer has no intention of exposing these
methods.

@@

\section{Conclusions and Future Work}
\label{sec:conclusion}
We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration
file for a specific JavaScript library. The tool downloads code examples written by the
developers from the library's repository. It uses these examples to execute the library and
gather data flow and type information. The tool finally generates a TypeScript declaration
file based on the information gathered at run time.

We developed an architecture that supports the automatic generation of declaration files
for specific JavaScript libraries without additional manual tasks. The architecture
contemplates future incorporation of a symbolic execution engine that refines the
initial code base enabling the exploration of new execution paths. Although it is not
implemented 
in this work, its incorporation would result in small incremental modifications to the
presented architecture as it is considered to only expand the existing code base.

Building an end-to-end solution for the generation of TypeScript declaration files was
prioritized over type inference accuracy. Consequently, types were taken over from the
values at run time. Developers express their intent how a library should be used with
example code in the documentation. Hence,
obtaining the types from the code examples extracted from the repositories proved to be a
pragmatic and effective approximation, enabling to work on specific aspects regarding the
TypeScript declaration file generation itself.

>>>>

\input{macros}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\coderef}[1]{Listing~\ref{#1}}

@@

  \caption{Generated declaration file for \texttt{glob-to-regexp}} 
  \label{fig:motivating-example-glob-to-regexp-vscode}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
  \caption[dts-generate - Architecture overview]{dts-generate - Architecture
    overview
    % - The initial code base is retrieved from the \NPM{} package's
    % repository.
    % Run-time information is gathered driven by example code.
    % A TypeScript declaration file is generated using run-time
    % information.
    % A Symbolic Execution Engine creates test cases based on the generated
    % Declaration File and via a feedback loop enriches the code base until the stopping
    % criteria is reached.
    % The final TypeScript Declaration File gets returned.
    % Feedback
    % loop through the Symbolic Execution Engine was not implemented. It can be added in a
    % future to the existing architecture, without modifying the working blocks.
  } 
  \label{fig:tsd_generation_method_block_diagram}
\end{figure}

@@

\begin{figure}[tp]
  \begin{centering}
      {\includegraphics[width=1\linewidth]{evaluation-diagram.pdf}}
      \caption[Evaluation against DefinitelyTyped repository]{{Evaluation of
          generated declaration files
          % against DefinitelyTyped          repository
        }
        % - The parser
        % uses the TypeScript Compiler           API \cite{typescript-compiler-api} to
        % transform declaration files into an abstract syntax which is serialized in
        % JSON. The comparison is performed on the JSON format.
} 
      \label{fig:evaluation-diagram}
  \end{centering}
\end{figure}

@@

\begin{lstlisting}[
  caption={Comparing declaration files without differences},
  language=bash,
label=code:dts-compare-example,
  float=tp,captionpos=b
]
$ dts-compare -e expected.d.ts -a actual.d.ts --module-name "my-module"
{
    "module": "my-module",
    "template": "module",
    "differences": [],
    "tags": []
}
\end{lstlisting}
% $

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=1\linewidth]{funnel.png}
  \caption[Number of analyzed modules for each stage of the experiment]{
    {Number of analyzed modules at each stage}
  % It was only possible to gather valid run-time information for about 25\% of the modules for which a Code Example was extracted.
  } 
  \label{fig:experiments-overall-funnel}
\end{figure}

@@

Next, we describe the results obtained for each step.

@@

\paragraph*{Generated Declaration Files}
A declaration file was generated for \CountModulesGeneratedDeclarationFile{} out of \CountModulesRunTimeInfoExtracted{} modules. Despite the correct execution
of the instrumented code examples, the extracted code for
\CountModulesDeclarationFileCouldNotBeGenerated{} modules did 
not exercise the library sufficiently. Hence, the collected run-time information was not enough
for generating a declaration file.

@@

\paragraph*{Literal types vs string}
Many JavaScript libraries use literal strings to select
options and configurations. To be concrete, 1925 DT packages rely on
literal types (see Table~\ref{tab:dts-parse-stats}). For example, the bonjour
package\footnote{\url{https://www.npmjs.com/package/@types/bonjour}}
relies on strings to control events and indicate protocols in (different) interfaces:
\begin{lstlisting}[language=TypeScript]
    removeAllListeners(event?: 'up' | 'down'): this;
    protocol?: 'udp'|'tcp';
\end{lstlisting}
The carlo
package\footnote{\url{https://www.npmjs.com/package/@types/carlo}}
uses literals to distinguish different events.
\begin{lstlisting}[language=TypeScript]
export type AppEvent = 'exit' | 'window';
\end{lstlisting}
% coinbase ...

@@

\section{Related Work}
\label{sec:related-work}
\paragraph*{Dynamic analysis techniques} Trace Typing
\cite{DBLP:conf/ecoop/AndreasenGCSTS16} is a framework for evaluating
retrofitted type systems. It involves gathering traces of JavaScript program
executions, just like in our system. However, the goal of the type
systems that they study is quite different from ours. They are
interested in type information to improve run-time performance and
traditional compilation, giving information about object layout and
tag tests. Contrary to their approach, we do not just observe the
types of arguments and return values. In addition, we also trace
property accesses which translate to requirements on objects. We
have no generalization step to infer polymorphic types.

Hummingbird \cite{DBLP:conf/pldi/RenF16} is a system for type checking
Ruby programs at run time in the presence of
metaprogramming. Metaprograms additionally generate
type annotations for the generated code. Once a generated method is
called, the method's body is statically checked against the actual
argument types as well as the generated annotations. This system does
not involve run-time tracing.

@@

JSTrace \cite{saftoiu21:_jstrac} performs dynamic type inference
for JavaScript by gathering information at run time. It inspects
run-time types of arguments and return values of suitably wrapped
functions. While the resulting information is treated similarly as in
our work, we gather interactions from property
accesses and method calls and derive further typing constraints from
those. Moreover, JSTrace performs an (expensive) deep inspection of
the types of objects and arrays whereas our inspection is shallow and
thus more efficient.

@@

TSInfer faces the problem
of including internal methods and private properties from the snapshot in the declaration file. Run-time
information would have shown that the developer has no intention of exposing these
methods.

@@

\section{Conclusions and Future Work}
\label{sec:conclusion}
We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration
file for a specific JavaScript library. It downloads the developer's
code examples from the library's repository. It uses these examples to
execute the library and gather data flow and type information. The tool generates a TypeScript declaration
file based on the information gathered at run time.

% We developed an architecture that supports the automatic generation of declaration files
% for specific JavaScript libraries without additional manual tasks. The architecture
% contemplates future incorporation of a symbolic execution engine that refines the
% initial code base enabling the exploration of new execution paths. Although it is not
% implemented 
% in this work, its incorporation would result in small incremental modifications to the
% presented architecture as it is considered to only expand the existing code base.

Building an end-to-end solution for the generation of TypeScript declaration files was
prioritized over type inference accuracy. Hence, types were taken over from the
values at run time. Developers express their intent how a library should be used with
example code in the documentation. 
Obtaining the types from the code examples extracted from the repositories proved to be a
pragmatic and effective approximation, enabling to work on specific aspects regarding the
TypeScript declaration file generation itself.
